---
title: "Stigler diet"
author: "Adam Sardar"
date: "1/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is heavily planned on a blog post by google OR tools

https://developers.google.com/optimization/lp/glop

https://en.wikipedia.org/wiki/Stigler_diet

https://pubsonline.informs.org/doi/pdf/10.1287/opre.49.1.1.11187

https://math.berkeley.edu/~mgu/MA170F2015/Diet.pdf


# Download Stigler Diert dataset from OR tools webpage

This saves me having to transcribe it from the original 1940's manuscript!

```{r}

library(rvest)
library(magrittr)
library(data.table)

StiglerDietData <- read_html("https://developers.google.com/optimization/lp/glop#data")
```

Extract out the nutrient RDAs.

```{r}
nutrientConstraintsTable <- StiglerDietData %>%
  html_nodes("table") %>%
  .[[2]] %>%
  html_table() %>%
  as.data.table()

setnames(nutrientConstraintsTable, c("Nutrient","RDI"))

nutrientConstraintsTable[, RDI_value := str_extract(RDI, "[\\d,\\.]+") %>% str_replace(",","") %>% as.numeric()]
nutrientConstraintsTable[, RDI_unit := str_extract(RDI, "\\w+$")]

nutrientConstraintsTable[Nutrient == "Vitamin A", RDI_value := RDI_value/1000] # Menu choices before are in 1000's of IU
```

And then the menu items, taking care to clean up the unusual data quantities

```{r}
menuChoicesTable <- StiglerDietData %>%
  html_nodes("table") %>%
  .[[3]] %>%
  html_table() %>% 
  as.data.table()

setnames(menuChoicesTable,
         c('Commodity', 'Weight', 'Price1939', nutrientConstraintsTable$Nutrient))

# Units are in thousands - see article
menuChoicesTable[, Calories := Calories*1000]
menuChoicesTable[, Price1939 := as.numeric(Price1939)]

# Weights taken from https://food.unl.edu/how-interpret-can-size-numbers
menuChoicesTable[Weight == "No. 2", Weight := "20 oz."]
menuChoicesTable[Weight == "No. 2 1/2", Weight := "29 oz."]

menuChoicesTable[Weight == "10 1/2", Weight := "10.5"]

menuChoicesTable[, Weight := str_replace_all(Weight, "1/2", "0.5")]
menuChoicesTable[, Weight := str_replace_all(Weight, "1/4", "0.25")]


menuChoicesTable[, Weight_unit := str_extract(Weight, "\\w+\\.?$") %>%
                                  str_remove("\\."), 
                 by = Commodity]

menuChoicesTable[, Weight_value := str_extract(Weight, "[\\d\\.]+") %>% as.numeric(),
                 by = Commodity]

setcolorder(menuChoicesTable, c('Commodity', 'Weight_unit', 'Weight_value'))

menuChoicesTable[,1:3]
```

Convert everything to oz (in cases where this is difficult, use nutrition per 100g from a regerence and map via calories).

```{r}
menuChoicesTable[Weight_unit == "oz"] # Do nothing

# 16 oz in a lb
menuChoicesTable[Weight_unit == "lb", Weight_value := Weight_value * 16]
menuChoicesTable[Weight_unit == "lb", Weight_unit := "oz"]

# 32 fluid oz in a quart, assume density of water
menuChoicesTable[Weight_unit == "qt", Weight_value := Weight_value * 32]
menuChoicesTable[Weight_unit == "qt", Weight_unit := "oz"]

# doz! 

#Calorific infor taken from google

menuChoicesTable[Weight_unit != "oz", 1:4]

menuChoicesTable[
  SJ( Commodity = 
    c("Eggs","Cream","Mayonnaise","Lemons","Oranges","Carrots","Celery","Lettuce"),
  CalPer100g = 
    c(155, 196, 680, 29, 47, 41, 18, 15)),
  Weight_value := round(Calories/(CalPer100g * 3.5),digits = 2),
  by = .EACHI, on = "Commodity"]


menuChoicesTable[Commodity %in% c("Eggs","Cream","Mayonnaise","Lemons","Oranges","Carrots","Celery","Lettuce"),
                 Weight_unit := "oz"]

################

menuChoicesTable[,unique(Weight_unit)] # Everythin is in oz

menuChoicesTable[,1:3]

```

Save dataset for use in package

```{r}

StiglerDiet <- list(nutrientRDA = nutrientConstraintsTable,
                    foodAvailable = menuChoicesTable)

save(StiglerDiet, file = "./data/StiglerDiet.RData", compress = "xz")
```



# Optimise for price

```{r}


costVec <- rep(1, length(menuChoicesTable$Commodity)) # everything will be per dollar spent
names(costVec) <- menuChoicesTable$Commodity

# Data is in amount per dollar
constraintMat <- menuChoicesTable[,as.matrix(.SD) ,
      .SDcols = !c("Commodity","Price1939","Weight","Weight_unit","Weight_value")]
row.names(constraintMat) <- menuChoicesTable$Commodity

objectiveVec <- nutrientConstraintsTable$RDI_value
names(objectiveVec) <- nutrientConstraintsTable$Nutrient

```


```{r}
refSol <-
  c(
  `Wheat Flour (Enriched)` = 0.029519,
  `Liver (Beef)` = 0.001893,
  Cabbage = 0.011214,
  Spinach = 0.005008,
  `Navy Beans, Dried` = 0.061029)

menuChoicesTable[, refProp := 0]

menuChoicesTable[SJ(Commodity = names(refSol),
                    proportion = refSol),
                refProp := proportion,
                 on = "Commodity"]

menuChoicesTable[refProp > 0]

```


```{r}
library(Rglpk)


opt <- Rglpk_solve_LP(
  obj = costVec,
  mat = t(constraintMat[names(costVec),]),
  dir = rep(">=", length(objectiveVec)),
  rhs = objectiveVec[colnames(constraintMat)],
  types = "C",
  max = FALSE
)



menuChoicesTable[, solProp := 0]
menuChoicesTable[SJ(Commodity = names(costVec),
                    proportion = opt$solution),
                solProp := proportion,
                 on = "Commodity"]

t(constraintMat[names(costVec),]) %*% menuChoicesTable$refProp

t(constraintMat[names(costVec),]) %*% menuChoicesTable$solProp

menuChoicesTable[solProp > 0]

menuChoicesTable[refProp > 0]



sols[ refProp > 0]

```



```{r}


microbenchmark::microbenchmark(

  glopt <- glop_lp(
    objectiveCoefs = costVec,
    constraintMat = t(constraintMat[names(costVec),]),
    constraintRHS = objectiveVec[colnames(constraintMat)]
  ),
  
opt <- Rglpk_solve_LP(
  obj = costVec,
  mat = t(constraintMat[names(costVec),]),
  dir = rep(">=", length(objectiveVec)),
  rhs = objectiveVec[colnames(constraintMat)],
  types = "C",
  max = FALSE
)

)
  

```



# Optimise for weight (off for a hike)


```{r}

weightPerDollar <- menuChoicesTable[,Weight_value/(Price1939/100)]
names(weightPerDollar) <- menuChoicesTable$Commodity

weightOpt <- Rglpk_solve_LP(
  obj = weightPerDollar,
  mat = t(constraintMat[names(weightPerDollar),]),
  dir = rep(">=", length(objectiveVec)),
  rhs = objectiveVec[colnames(constraintMat)],
  types = "C",
  max = FALSE
)

menuChoicesTable[SJ(Commodity = names(weightPerDollar),
                    proportion = weightOpt$solution),
                weightSol := proportion,
                 on = "Commodity"][weightSol > 0]

t(constraintMat[names(costVec),]) %*% menuChoicesTable$weightSol

```


# As calories change


Use a larger dataset and current RDA

https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/methods-and-application-of-food-composition-laboratory/mafcl-site-pages/sr17-sr28/


# What about carbon?





